% !TEX root = ../main.tex
This paper presents a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling. We introduce an adaptation of the Anderson-Darling test, which we name the \textit{marginalized Anderson-Darling test}, to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.

The statistical approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation, and better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.

We do not see a relationship between bar strength and pitch angle in our data, as would have been predicted by Manifold theory, and do not find evidence for the relationship between bulge size and pitch angle predicted by the Hubble sequence. \citet{2007ApJ...655...77G} propose that bulge S\'ersic index is correlated with black hole mass, which has in turn been shown to correlate with spiral pitch angle, we do not investigate this here as \textit{Galaxy Builder} could not accurately recover bulge S\'ersic indices.

Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and reccurent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time. However, the assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$? Is the spiral arm equally apparent at all pitch angles, or is a selection effect present? This result is also not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.

We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles within the limits specified above. We also assume that the galaxy sample is representative of the general spiral population, which is not necessarily the case. As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.

The methodlogy proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
